---
title: "Soapbox event Kamloops 2024"
author: "Jessica Alligham"
date: "2024-08-29"
output: html_document
---


```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(stringr)
library(readxl) # Use readr::read_csv if the file is in CSV format

# Step 1: Load the data (assuming the file is an Excel file)

data <- read_excel("Soapbox Science Community Outreach Impact_34observations.xlsx", sheet = 1)

# Step 2: Remove unnecessary columns
columns_to_remove <- c(
  "Respondent ID", "Collector ID", "Start Date", "End Date", "IP Address", 
  "Email Address", "First Name", "Last Name", "Custom Data 1", 
  "Do you consent to having your data in the study?", 
  "What is your identifying number which will only be used if you wish to withdraw your data from the study after completion?",
  "Thank you for participating in our study and taking part in Kamloops’ Soapbox Science event."
)

# Remove the unnecessary columns
data_cleaned <- data %>% 
  dplyr::select(-all_of(columns_to_remove))  
data_cleaned<-data_cleaned[-1,-1]

columns_to_merge <- c("Science and research improve our lives.", "...14", "...15", "...16")

# Step 3: Merge the columns into a single column
# We use coalesce() which will take the first non-NA value from the specified columns for each row
data_cleaned_merge1 <- data_cleaned %>%
  mutate(`Science and research improve our lives.` = coalesce(!!!syms(columns_to_merge)))

# Step 4: Remove the original option columns
data_cleaned_merge1 <- data_cleaned_merge1 %>% 
  dplyr::select(-all_of(columns_to_merge[2:4]))

# Step 5: View the cleaned data to verify
names(data_cleaned_merge1)
head(data_cleaned_merge1)
#View(data_cleaned_merge1)# Display the first few rows to verify
```

```{r}
# Step 6: Merge the next set of columns for "I am good with observation and problems solving (the scientific method)."
columns_to_merge2 <- c("I am good with observation and problems solving (the scientific method).", "...18", "...19", "...20")

data_cleaned_merge2 <- data_cleaned_merge1 %>%
  mutate(`I am good with observation and problems solving (the scientific method).` = coalesce(!!!syms(columns_to_merge2)))

# Remove the original option columns for the second question
data_cleaned_merge2 <- data_cleaned_merge2 %>% 
  dplyr::select(-all_of(columns_to_merge2[2:4]))


names(data_cleaned_merge2)
head(data_cleaned_merge2)
#View(data_cleaned_merge2) # Display the first few rows to verify
```


```{r}
# Step 7: Merge the next set of columns for "I would consider starting my career in or switching my career to be in the field of science, technology, engineering, or math if given the opportunity to do so."
columns_to_merge3 <- c("I would consider starting my career in or switching my career to be in the field of science, technology, engineering, or math if given the opportunity to do so.", "...22", "...23", "...24")

data_cleaned_merge3 <- data_cleaned_merge2 %>%
  mutate(`I would consider starting my career in or switching my career to be in the field of science, technology, engineering, or math if given the opportunity to do so.` = coalesce(!!!syms(columns_to_merge3)))

# Remove the original option columns for the second question
data_cleaned_merge3 <- data_cleaned_merge3 %>% 
  dplyr::select(-all_of(columns_to_merge3[2:4]))


names(data_cleaned_merge3)
head(data_cleaned_merge3)
#View(data_cleaned_merge3) # Display the first few rows to verify
```

```{r}
# Step 8: Merge the next set of columns for "It was inspirational to see women and non-binary scientists sharing their research."
columns_to_merge4 <- c("It was inspirational to see women and non-binary scientists sharing their research.", "...26", "...27", "...28")

data_cleaned_merge4 <- data_cleaned_merge3 %>%
  mutate(`It was inspirational to see women and non-binary scientists sharing their research.` = coalesce(!!!syms(columns_to_merge4)))

# Remove the original option columns 
data_cleaned_merge4 <- data_cleaned_merge4 %>% 
  dplyr::select(-all_of(columns_to_merge4[2:4]))


names(data_cleaned_merge4)
head(data_cleaned_merge4)
#View(data_cleaned_merge4) # Display the first few rows to verify
```

```{r}
# Step 9: Merge the next set of columns for "The Kamloops Soapbox Science event increased my interest in science and research."
columns_to_merge5 <- c("The Kamloops Soapbox Science event increased my interest in science and research.", "...30", "...31", "...32")

data_cleaned_merge5 <- data_cleaned_merge4 %>%
  mutate(`The Kamloops Soapbox Science event increased my interest in science and research.` = coalesce(!!!syms(columns_to_merge5)))

# Remove the original option columns for the second question
data_cleaned_merge5 <- data_cleaned_merge5 %>% 
  dplyr::select(-all_of(columns_to_merge5[2:4]))


names(data_cleaned_merge5)
head(data_cleaned_merge5)
#View(data_cleaned_merge5) # Display the first few rows to verify
```

```{r}
# Step 10: Merge the next set of columns for "How interested are you in science and research?"
columns_to_merge6 <- c("How interested are you inscience and research?", "...34", "...35", "...36")

data_cleaned_merge6 <- data_cleaned_merge5 %>%
  mutate(`How interested are you inscience and research?` = coalesce(!!!syms(columns_to_merge6)))

# Remove the original option columns
data_cleaned_merge6 <- data_cleaned_merge6 %>% 
  dplyr::select(-all_of(columns_to_merge6[2:4]))


names(data_cleaned_merge6)
head(data_cleaned_merge6)
View(data_cleaned_merge6) # Display the first few rows to verify
```



```{r}
# Step 11: Merge the next set of columns for "What is your educational background?"
columns_to_merge7 <- c("What is your educational background?", "...78", "...79","...80","...81","...82","...83")

data_cleaned_merge7 <- data_cleaned_merge6 %>%
  mutate(`What is your educational background?` = coalesce(!!!syms(columns_to_merge7)))

# Remove the original option columns
data_cleaned_merge7 <- data_cleaned_merge7 %>% 
  dplyr::select(-all_of(columns_to_merge7[2:7]))


names(data_cleaned_merge7)
head(data_cleaned_merge7)
#View(data_cleaned_merge7) # Display the first few rows to verify
```

```{r}
# Step 12: Merge the next set of columns for "What is your educational background?"
columns_to_merge8 <- c("What is your gender?", "...58", "...59","...60","...61","...62","...63","...64","...65")

data_cleaned_merge8 <- data_cleaned_merge7 %>%
  mutate(`What is your gender?` = coalesce(!!!syms(columns_to_merge8)))

# Remove the original option columns
data_cleaned_merge8 <- data_cleaned_merge8 %>% 
  dplyr::select(-all_of(columns_to_merge8[2:9]))


names(data_cleaned_merge8)
head(data_cleaned_merge8)
View(data_cleaned_merge8) # Display the first few rows to verify
```

```{r}
# Step 13: Merge the next set of columns for "What is your educational background?"
columns_to_merge9 <- c("How old are you?", "...50", "...51","...52","...53","...54","...55","...56")

data_cleaned_merge9 <- data_cleaned_merge8 %>%
  mutate(`How old are you?` = coalesce(!!!syms(columns_to_merge9)))

# Remove the original option columns
data_cleaned_merge9 <- data_cleaned_merge9 %>% 
  dplyr::select(-all_of(columns_to_merge9[2:8]))


names(data_cleaned_merge9)
head(data_cleaned_merge9)
#View(data_cleaned_merge9) # Display the first few rows to verify
```

```{r}

# Step 14: Split and Extract Features for Each Column
# Columns to process for multiple options per respondent
columns_to_process10 <- c("Who below looks like a scientist (select all that apply)?", "...38", "...39", "...40", "...41", 
                          "...42", "...43", "...44", "...45", "...46", "...47", "...48")


data_cleaned_split <- data_cleaned_merge9  # Initialize the new DataFrame

for (i in seq_along(columns_to_process10)) {
  # Column name to process
  column_name <- columns_to_process10[i]
  
  # Extract features for each selection in the specified column
  data_cleaned_split <- data_cleaned_split %>%
    mutate(
      !!paste0("Age_", i) := ifelse(str_detect(!!sym(column_name), "middle aged"), "Middle aged",
                                    ifelse(str_detect(!!sym(column_name), "older"), "Older", NA)),
      
      # Improved Ethnicity detection with better logic to handle "white" in phrases
      !!paste0("Ethnicity_", i) := ifelse(
        str_detect(!!sym(column_name), "white"), "White",
        ifelse(str_detect(!!sym(column_name), "black"), "Black",
               ifelse(str_detect(!!sym(column_name), "asian"), "Asian", NA))),
      
      # Ensure 'female' is checked before 'male'
      !!paste0("Gender_", i) := ifelse(str_detect(!!sym(column_name), "female"), "Female",
                                       ifelse(str_detect(!!sym(column_name), "male"), "Male", NA)),
      
      # Consolidate Dressing into "Casual" and "Formal"
      !!paste0("Dressing_", i) := case_when(
        str_detect(!!sym(column_name), "red shirt|green shirt|red tank top|graphic t-shirt| headband| tank top|backwards ball cap|ball cap") ~ "Casual",
        str_detect(!!sym(column_name), "pink shirt|lab coat|red jacket") ~ "Formal",
        TRUE ~ NA
      ),
      
      # Consolidate Expression into "Positive," "Neutral," and "Serious"
      !!paste0("Expression_", i) := case_when(
        str_detect(!!sym(column_name), "smiling|big smile") ~ "Positive",
        str_detect(!!sym(column_name), "straight face") ~ "Neutral",
        TRUE ~ NA
      )
    )
}

# View the split data to verify
#View(data_cleaned_split)

# Step 2: Remove the Original Columns After Splitting
data_cleaned_after_split <- data_cleaned_split %>%
  dplyr::select(-all_of(columns_to_process10))  # Remove the original columns

# View the data after removing original columns
#View(data_cleaned_after_split)

# Step 3: Create Binary Indicators for Each Category
binary_columns <- list(
  "Middle aged" = "Scientist_Age_MiddleAged",
  "Older" = "Scientist_Age_Older",
  "White" = "Scientist_Ethnicity_White",
  "Black" = "Scientist_Ethnicity_Black",
  "Asian" = "Scientist_Ethnicity_Asian",
  "Male" = "Scientist_Gender_Male",
  "Female" = "Scientist_Gender_Female",
  "Casual" = "Scientist_Dressing_Casual",
  "Formal" = "Scientist_Dressing_Formal",
  "Positive" = "Scientist_Expression_Positive",
  "Neutral" = "Scientist_Expression_Neutral"
)

# Initialize binary columns
for (binary_name in binary_columns) {
  data_cleaned_after_split[[binary_name]] <- 0
}

# Assign binary values based on existing categories
for (binary_col in names(binary_columns)) {
  target_col <- binary_columns[[binary_col]]
  data_cleaned_after_split[[target_col]] <- apply(data_cleaned_after_split, 1, function(row) {
    ifelse(any(row == binary_col, na.rm = TRUE), 1, 0)
  })
}

# View the data with binary indicators
#View(data_cleaned_after_split)

#  Remove the Intermediate Split Columns After Consolidation
data_merged10 <- data_cleaned_after_split %>%
  dplyr::select(-matches("Age_\\d+|Ethnicity_\\d+|Gender_\\d+|Dressing_\\d+|Expression_\\d+"))  # Remove intermediate split columns

# View the final consolidated data
#View(data_merged10)

```


```{r}


# Columns to process for multiple options per respondent (ethnic group question)
columns_to_process_ethnic <- c("To which ethnic group do you belong, select all that apply?", "...67", "...68", "...69", "...70", 
                          "...71", "...72", "...73", "...74", "...75", "...76")

# Ethnic group options
ethnic_options <- list(
  "Black" = "Response_Ethnicity_Black",
  "White" = "Response_Ethnicity_White",
  "East Asian" = "Response_Ethnicity_EastAsian",
  "South Asian" = "Response_Ethnicity_SouthAsian",
  "Southeast Asian" = "Response_Ethnicity_SoutheastAsian",
  "Latino" = "Response_Ethnicity_Latino",
  "Indigenous" = "Response_Ethnicity_Indigenous",
  "Middle Eastern / West Asian" = "Response_Ethnicity_MiddleEastern_WestAsian",
  "Not listed" = "Response_Ethnicity_NotListed",
  "I am not certain" = "Response_Ethnicity_NotCertain",
  "Prefer not to answer" = "Response_Ethnicity_PreferNotToAnswer"
)

# Step 15: Split and Extract Ethnic Group Features for Each Column
data_cleaned_split_ethnic <- data_merged10  # Use the final cleaned data from the previous steps

for (i in seq_along(columns_to_process_ethnic)) {
  # Column name to process
  column_name <- columns_to_process_ethnic[i]
  
  # Extract features for each selection in the specified column
  data_cleaned_split_ethnic <- data_cleaned_split_ethnic %>%
    mutate(
      # Binary indicators for each ethnicity option
      !!paste0("Ethnicity_Black_", i) := ifelse(str_detect(!!sym(column_name), "Black"), 1, 0),
      !!paste0("Ethnicity_White_", i) := ifelse(str_detect(!!sym(column_name), "White"), 1, 0),
      !!paste0("Ethnicity_EastAsian_", i) := ifelse(str_detect(!!sym(column_name), "East Asian"), 1, 0),
      !!paste0("Ethnicity_SouthAsian_", i) := ifelse(str_detect(!!sym(column_name), "South Asian"), 1, 0),
      !!paste0("Ethnicity_SoutheastAsian_", i) := ifelse(str_detect(!!sym(column_name), "Southeast Asian"), 1, 0),
      !!paste0("Ethnicity_Latino_", i) := ifelse(str_detect(!!sym(column_name), "Latino"), 1, 0),
      !!paste0("Ethnicity_Indigenous_", i) := ifelse(str_detect(!!sym(column_name), "Indigenous"), 1, 0),
      !!paste0("Ethnicity_MiddleEastern_WestAsian_", i) := ifelse(str_detect(!!sym(column_name), "Middle Eastern / West Asian"), 1, 0),
      !!paste0("Ethnicity_NotListed_", i) := ifelse(str_detect(!!sym(column_name), "Not listed"), 1, 0),
      !!paste0("Ethnicity_NotCertain_", i) := ifelse(str_detect(!!sym(column_name), "I am not certain"), 1, 0),
      !!paste0("Ethnicity_PreferNotToAnswer_", i) := ifelse(str_detect(!!sym(column_name), "Prefer not to answer"), 1, 0)
    )
}

# View the split data to verify
#View(data_cleaned_split_ethnic)

# Remove the Original Columns After Splitting
data_cleaned_after_split_ethnic <- data_cleaned_split_ethnic %>%
  dplyr::select(-all_of(columns_to_process_ethnic))  # Remove the original columns

# View the data after removing original columns
#View(data_cleaned_after_split_ethnic)

# Consolidate Binary Indicators for Each Ethnic Group
for (binary_col in names(ethnic_options)) {
  target_col <- ethnic_options[[binary_col]]
  ethnic_pattern <- paste0("Ethnicity_", gsub("Response_Ethnicity_", "", target_col), "_")
  
  # Select columns matching the pattern using grep
  relevant_columns <- grep(ethnic_pattern, names(data_cleaned_after_split_ethnic), value = TRUE)
  
  # Check across all the new split columns and set consolidated value
  data_cleaned_after_split_ethnic[[target_col]] <- apply(data_cleaned_after_split_ethnic[, relevant_columns], 1, function(row) {
    ifelse(any(row == 1, na.rm = TRUE), 1, 0)
  })
}

# View the consolidated data to verify
#View(data_cleaned_after_split_ethnic)

# Remove the Intermediate Split Columns After Consolidation
final_data_cleaned <- data_cleaned_after_split_ethnic %>%
  dplyr::select(-matches("Ethnicity_\\w+_\\d+"))  # Remove intermediate split columns

# View the final consolidated data
View(final_data_cleaned)

```


```{r}
#install.packages("openxlsx")
library(openxlsx)

write.xlsx(final_data_cleaned, "final_data_cleaned.xlsx")


```


```{r}
# Read the cleaned CSV file into R
soapbox_data <-  read_excel("final_data_cleaned.xlsx")

# View the first few rows of the data to verify
head(soapbox_data)
View(soapbox_data)
names(soapbox_data)

```




```{r}


# Rename columns with meaningful names
colnames(soapbox_data) <- c(
  "Improves_Lives",                      # Science and research improve our lives.
  "Good_With_Scientific_Method",         # I am good with observation and problem solving (the scientific method).
  "Consider_Career_In_STEM",             # I would consider starting my career or switching to a career in STEM.
  "Inspired_By_Women_NonBinary_Scientists", # It was inspirational to see women and non-binary scientists sharing their research.
  "Interest_Increased_By_Event",         # The Kamloops Soapbox Science event increased my interest in science and research.
  "Interest_In_Science_Research",        # How interested are you in science and research?
  "Age",                                 # How old are you?
  "Gender",                              # What is your gender?
  "Educational_Background",              # What is your educational background?
  "Scientist_Age_MiddleAged",            # Middle-aged indicator
  "Scientist_Age_Older",                 # Older indicator
  "Scientist_Ethnicity_White",           # White Ethnicity indicator for scientist
  "Scientist_Ethnicity_Black",           # Black Ethnicity indicator for scientist
  "Scientist_Ethnicity_Asian",           # Asian Ethnicity indicator for scientist
  "Scientist_Gender_Male",               # Male Gender indicator for scientist
  "Scientist_Gender_Female",             # Female Gender indicator for scientist
  "Scientist_Dressing_Casual",           # Casual Dressing indicator for scientist
  "Scientist_Dressing_Formal",           # Formal Dressing indicator for scientist
  "Scientist_Expression_Positive",       # Positive Expression indicator for scientist
  "Scientist_Expression_Neutral",        # Neutral Expression indicator for scientist
  "Response_Ethnicity_Black",            # Respondent Ethnicity: Black
  "Response_Ethnicity_White",            # Respondent Ethnicity: White
  "Response_Ethnicity_EastAsian",        # Respondent Ethnicity: East Asian
  "Response_Ethnicity_SouthAsian",       # Respondent Ethnicity: South Asian
  "Response_Ethnicity_SoutheastAsian",   # Respondent Ethnicity: Southeast Asian
  "Response_Ethnicity_Latino",           # Respondent Ethnicity: Latino
  "Response_Ethnicity_Indigenous",       # Respondent Ethnicity: Indigenous
  "Response_Ethnicity_MiddleEastern",    # Respondent Ethnicity: Middle Eastern / West Asian
  "Response_Ethnicity_NotListed",        # Respondent Ethnicity: Not listed
  "Response_Ethnicity_NotCertain",       # Respondent Ethnicity: Not certain
  "Response_Ethnicity_PreferNotToAnswer" # Respondent Ethnicity: Prefer not to answer
)

# View the renamed data frame
head(soapbox_data)
View(soapbox_data)
names(soapbox_data)
```

```{r}

soapbox_data<-na.omit(soapbox_data)
sum(is.na(soapbox_data))
```



```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Create a mapping of column names to shorter titles
short_titles <- c(
  "Improves_Lives" = "Science improves lives",
  "Good_With_Scientific_Method" = "Proficiency in science",
  "Consider_Career_In_STEM" = "Considering a STEM career",
  "Inspired_By_Women_NonBinary_Scientists" = "Inspired by scientists",
  "Interest_Increased_By_Event" = "Interest increased by event",
  "Interest_In_Science_Research" = "Interest in research",
  "Age" = "Age distribution",
  "Gender" = "Gender distribution",
  "Educational_Background" = "Educational background"
)

# Define a function to create a bar plot with dynamic label positioning
plot_categorical <- function(data, column_name, plot_title) {
  # Calculate counts and percentages
  plot_data <- data %>%
    group_by(.data[[column_name]]) %>%
    summarise(Count = n(), .groups = "drop") %>%
    mutate(Percentage = round((Count / sum(Count)) * 100, 1),
           Label = paste0(Count, " (", Percentage, "%)"))  # Labels for bars
  
  # Create the horizontal bar plot
  ggplot(plot_data, aes(x = Count, y = reorder(.data[[column_name]], Count), fill = Count)) +
    geom_bar(stat = "identity", color = "black", width = 0.8) +  # Bars with black borders
    
    # Dynamic label placement: Inside for wide bars, outside for narrow bars
    geom_text(aes(label = Label,
                  hjust = ifelse(Count / max(Count) > 0.3, 1.1, -0.1)),  # Inside or outside based on bar size
              color = ifelse(plot_data$Count / max(plot_data$Count) > 0.3, "white", "black"), 
              size = 5, fontface = "bold") +
    
    scale_fill_gradient(low = "#D6EAF8", high = "#1B4F72") +  # Gradient fill
    theme_minimal(base_size = 18) +  # Larger font size for readability
    labs(title = plot_title, x = "Count", y = NULL) +  # Title and x-axis label
    theme(
      legend.position = "none",  # Remove legend
      plot.title = element_text(size = 20, face = "bold", hjust = 0.5),  # Centered title
      axis.title.x = element_text(size = 16, face = "bold"),
      axis.text.x = element_text(size = 14, color = "black", face = "bold"),
      axis.text.y = element_text(size = 16, color = "black", face = "bold"),
      panel.grid.major = element_line(color = "lightgray"),
      panel.grid.minor = element_blank()
    ) +
    xlim(0, max(plot_data$Count) * 1.2)  # Add padding for labels outside the bars
}

# Load your dataset (ensure `soapbox_data` is available)
# Example: soapbox_data <- read.csv("your_data.csv")

# Generate and save individual plots with improved readability
for (col in names(short_titles)) {
  # Define file name dynamically
  file_name <- paste0(col, "_plot.png")
  
  # Save the plot to PNG
  png(filename = file_name, width = 1200, height = 900, res = 150)
  print(plot_categorical(soapbox_data, col, short_titles[[col]]))
  dev.off()
  
  cat("Saved:", file_name, "\n")  # Confirmation message for each saved file
}

```

```{r,eval=FALSE}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Create a mapping of column names to shorter titles
short_titles <- c(
  "Improves_Lives" = "Science improves lives",
  "Good_With_Scientific_Method" = "Proficiency in science",
  "Consider_Career_In_STEM" = "Considering a STEM career",
  "Inspired_By_Women_NonBinary_Scientists" = "Inspired by scientists",
  "Interest_Increased_By_Event" = "Interest increased by event",
  "Interest_In_Science_Research" = "Interest in research",
  "Age" = "Age distribution",
  "Gender" = "Gender distribution",
  "Educational_Background" = "Educational background"
)

# Define a function to create a bar plot with enhanced readability
plot_categorical <- function(data, column_name, plot_title) {
  # Calculate counts and percentages
  plot_data <- data %>%
    group_by(.data[[column_name]]) %>%
    summarise(Count = n(), .groups = "drop") %>%
    mutate(Percentage = round((Count / sum(Count)) * 100, 1),
           Label = paste0(Count, " (", Percentage, "%)"))  # Labels for inside bars
  
  # Create the horizontal bar plot
  ggplot(plot_data, aes(x = Count, y = reorder(.data[[column_name]], Count), fill = Count)) +
    geom_bar(stat = "identity", color = "black", width = 0.8) +  # Increase bar width
    geom_text(aes(label = Label), 
              hjust = 1.1, color = "white", size = 5, fontface = "bold") +  # Inside bar labels
    scale_fill_gradient(low = "#D6EAF8", high = "#1B4F72") +  # Gradient from light to dark blue
    theme_minimal(base_size = 18) +  # Larger base font size
    labs(title = plot_title, x = "Count", y = NULL) +  # Add clean title and x-axis label
    theme(
      legend.position = "none",  # Remove legend
      plot.title = element_text(size = 20, face = "bold", hjust = 0.5),  # Larger, centered title
      axis.title.x = element_text(size = 16, face = "bold"),  # Larger x-axis label
      axis.text.x = element_text(size = 14, color = "black", face = "bold"),  # X-axis text
      axis.text.y = element_text(size = 16, color = "black", face = "bold"),  # Y-axis text
      panel.grid.major = element_line(color = "lightgray"),  # Subtle grid lines
      panel.grid.minor = element_blank()  # Remove minor grid lines
    ) +
    xlim(0, max(plot_data$Count) * 1.2)  # Add padding for labels
}

# Load your dataset (ensure `soapbox_data` is available)
# Example: soapbox_data <- read.csv("your_data.csv")

# Generate and save individual plots
for (col in names(short_titles)) {
  # Define file name dynamically
  file_name <- paste0(col, "_plot.png")
  
  # Save the plot to PNG
  png(filename = file_name, width = 1200, height = 900, res = 150)
  print(plot_categorical(soapbox_data, col, short_titles[[col]]))
  dev.off()
  
  cat("Saved:", file_name, "\n")  # Confirmation message for each saved file
}

```




```{r,eval=FALSE}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Create a mapping of column names to shorter titles
short_titles <- c(
  "Improves_Lives" = "Science improves lives",
  "Good_With_Scientific_Method" = "Proficiency in science",
  "Consider_Career_In_STEM" = "Considering a STEM career",
  "Inspired_By_Women_NonBinary_Scientists" = "Inspired by scientists",
  "Interest_Increased_By_Event" = "Interest increased by event",
  "Interest_In_Science_Research" = "Interest in research",
  "Age" = "Age distribution",
  "Gender" = "Gender distribution",
  "Educational_Background" = "Educational background"
)

# Define the plotting function with shorter titles and reduced title size
plot_categorical <- function(data, column_name, plot_title) {
  # Calculate counts and percentages
  plot_data <- data %>%
    group_by(.data[[column_name]]) %>%
    summarise(Count = n(), .groups = "drop") %>%
    mutate(Percentage = round((Count / sum(Count)) * 100, 1))  # Calculate percentage
  
  # Create the horizontal bar plot
  ggplot(plot_data, aes(x = Count, y = reorder(.data[[column_name]], Count), fill = Count)) +
    geom_bar(stat = 'identity', color = "black") +  # Gradient fill with black border
    geom_text(aes(label = paste0(Count, " (", Percentage, "%)")), 
              hjust = -0.1, color = "black", size = 5, fontface = "bold") +  # Adjust text label size
    scale_fill_gradient(low = "lightblue", high = "darkblue") +  # Gradient color from light to dark blue
    theme_minimal(base_size = 14, base_family = "Helvetica") +  # Minimal white background
    labs(title = plot_title, x = "Count", y = NULL) +  # Use the explicit title
    theme(
      legend.position = "none",  # Remove the legend
      plot.title = element_text(size = 14, face = "bold"),  # Reduce title size
      axis.title.x = element_text(size = 14, face = "bold"),  # Adjust x-axis label size
      axis.text.x = element_text(size = 14, color = "black", face = "bold"),  # Adjust x-axis text size
      axis.text.y = element_text(size = 14, color = "black", face = "bold"),  # Adjust y-axis text size
      plot.background = element_rect(fill = "white", color = NA),  # White background
      panel.background = element_rect(fill = "white", color = NA),  # White panel background
      panel.grid.major = element_line(color = "gray"),  # Gray grid lines
      panel.grid.minor = element_blank()  # No minor grid lines
    ) +
    xlim(0, max(plot_data$Count) * 1.1)  # Adjust x-axis limits to ensure labels fit inside the plot
}

# Generate individual plots with shorter titles
for (col in names(short_titles)) {
  png(paste0(col, ".png"), width = 1200, height = 900, res = 150)  # Save each plot as a PNG
  print(plot_categorical(soapbox_data, col, short_titles[[col]]))
  dev.off()
}

```

```{r}


# List of categorical columns
categorical_columns <- c("Improves_Lives", "Good_With_Scientific_Method", 
                         "Consider_Career_In_STEM", "Inspired_By_Women_NonBinary_Scientists",
                         "Interest_Increased_By_Event", "Interest_In_Science_Research",
                         "Age", "Gender", "Educational_Background")


# Save summary tables as CSV
summary_tables <- lapply(categorical_columns, function(col) {
  soapbox_data %>%
    group_by(.data[[col]]) %>%
    summarise(Count = n(), .groups = "drop") %>%
    mutate(Percentage = round((Count / sum(Count)) * 100, 1)) %>%
    mutate(Variable = col)
})
combined_summary <- bind_rows(summary_tables)
write.csv(combined_summary, "Categorical_Variables_Summary.csv", row.names = FALSE)



```



The descriptive statistics provide valuable insights into participants' perceptions and backgrounds regarding science, research, and STEM careers. The analysis of each variable is presented below:

Improves Lives:
A vast majority of participants (93.3%) agreed that science and research improve lives, with 28 out of 30 respondents endorsing this view. Only 6.7% disagreed, indicating a strong consensus on the positive impact of science and research on life improvement.

Good With Scientific Method:
Most participants (63.3%, or 19 out of 30) reported confidence in their abilities related to the scientific method. Smaller proportions expressed varied levels of agreement or disagreement: 26.7% somewhat agreed, 6.7% completely disagreed, and 3.3% somewhat disagreed. This distribution suggests that while the majority feel confident in their scientific skills, there is a range of confidence levels among participants.

Consider Career in STEM:
The responses regarding motivation to pursue a career in STEM were spread across different levels of agreement. About 43.3% of participants were highly motivated to consider a STEM career, while the remaining responses indicated lower levels of interest. This varied distribution points to mixed feelings among participants about pursuing careers in STEM fields, with a significant portion demonstrating strong motivation.

Inspired by Women and Non-Binary Scientists:
Most participants (93.3%) reported being inspired by women and non-binary scientists, highlighting strong support for diverse role models in science. Only 6.7% somewhat agreed, suggesting that diverse representation in science is highly impactful for the majority of participants.

Interest Increased by Event:
A significant proportion of participants (70%) indicated that their interest in science and research increased due to a specific event, while 30% reported somewhat agrees about the increases in interest. These findings suggest that events can effectively boost interest in science for most participants, although a notable group remains less affected.

Interest in Science Research:
Interest in science research varied among participants, with 56.7% expressing high interest, 26.7% indicating medium interest, and 16.7% reporting low interest. Overall, these responses reflect a generally positive attitude towards science research, with over half of the respondents showing strong interest.

Age:
The age distribution of participants is spread across several categories, with the highest percentage (40%) falling within a specific age group of 31-40. Moreover,most of the participants were from age group 21-50 with 76.7%.Smaller percentages are observed in other age categories along with 2 participants prefered not to ans , indicating diversity in the age range of the sample, which could potentially influence the overall results.

Gender:
The majority of participants (60%) who responded were females followed by 33.3% males, while 1 participant from non binary gender and 1 prefer not to ans. This distribution may influence the interpretation of gender-related analyses due to the predominance of one group.

Educational Background:
Participants' educational backgrounds vary, with the highest percentage (23.3%) from undergrad and grad each followed by high school of 20% students. Other levels of education are represented in smaller proportions college diploma of 13.3% and while red seal and elementary schol with 10%, suggesting different levels of scientific literacy and experience among participants. This variation could affect their responses to other variables.

Overall Conclusion
The results indicate strong support for the positive impact of science, the influence of diverse role models, and the interest in STEM careers. However, the diversity in responses across variables such as career consideration, confidence in scientific methods, and educational background underscores the need for targeted strategies to address different segments within the population. This approach will help to better understand and support the varying levels of engagement and motivation among different groups.





```{r,eval=FALSE}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(tibble)

# Set up the data preparation
prepare_summary_df <- function(data, attribute_columns, total_participants) {
  # Calculate frequency of each attribute
  summary_stats <- colSums(data[, attribute_columns])
  
  # Convert summary stats to a data frame
  summary_df <- as.data.frame(summary_stats) %>%
    rownames_to_column(var = "Attribute") %>%
    rename(Frequency = summary_stats) %>%
    mutate(Percentage = round((Frequency / total_participants) * 100, 1))  # Calculate percentage for each attribute
  
  return(summary_df)
}

# Plotting function for the summary statistics with enhanced text readability
plot_summary <- function(summary_df) {
  ggplot(summary_df, aes(x = Attribute, y = Frequency, fill = Attribute)) +
    geom_bar(stat = "identity", fill = "blue") +  # Set bars to blue
    geom_text(aes(label = paste(Frequency, " (", Percentage, "%)", sep = "")), 
              vjust = -0.5, color = "black", size = 6) +  # Larger text labels for counts and percentages
    theme_minimal(base_size = 16, base_family = "Helvetica") +  # Increase base font size for readability
    labs(title = "Distribution of Scientist Attributes", x = NULL, y = "Frequency") +
    theme(
      legend.position = "none",  # Remove the legend
      axis.text.x = element_text(angle = 30, hjust = 1, color = "black", size = 14, face = "bold"),  # Slant x-axis text at 30 degrees, increase size and bold
      axis.text.y = element_text(color = "black", size = 14, face = "bold"),  # Ensure y-axis text is black and bold
      plot.title = element_text(size = 20, hjust = 0.5, face = "bold"),  # Centered and larger title
      plot.background = element_rect(fill = "white", color = NA),  # White background
      panel.background = element_rect(fill = "white", color = NA)  # White background for the panel
    )
}

# Assuming `soapbox_data` is the data frame and contains the necessary data
# Replace this with the actual number of participants
total_participants <- nrow(soapbox_data)

# List of attribute columns for scientist characteristics
attribute_columns <- c("Scientist_Age_MiddleAged", "Scientist_Age_Older", 
                       "Scientist_Ethnicity_White", "Scientist_Ethnicity_Black", 
                       "Scientist_Ethnicity_Asian", "Scientist_Gender_Male", 
                       "Scientist_Gender_Female", "Scientist_Dressing_Casual", 
                       "Scientist_Dressing_Formal", "Scientist_Expression_Positive", 
                       "Scientist_Expression_Neutral")

# Prepare the summary data frame
summary_df <- prepare_summary_df(soapbox_data, attribute_columns, total_participants)

# Display the table
print(summary_df)

# Find the attribute chosen the least
least_chosen_attribute <- summary_df %>% 
  filter(Frequency == min(Frequency)) %>%
  select(Attribute)
cat("The attribute chosen the least is:", least_chosen_attribute$Attribute, "\n")

# Save the plot as a PNG file with increased resolution
png("Scientist_Attributes_Distribution.png", width = 1200, height = 900, res = 150)  # Adjusted dimensions for better readability
plot_summary(summary_df)  # Generate the plot
dev.off()  # Close the device


```



```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(tibble)

# Data preparation function: Summarizes attribute frequencies and calculates percentages
prepare_summary_df <- function(data, attribute_columns, total_participants) {
  # Calculate total frequency of each attribute
  summary_stats <- colSums(data[, attribute_columns], na.rm = TRUE)
  
  # Convert summary stats to a clean data frame
  summary_df <- as.data.frame(summary_stats) %>%
    rownames_to_column(var = "Attribute") %>%
    rename(Frequency = summary_stats) %>%
    mutate(Percentage = round((Frequency / total_participants) * 100, 1),
           Label = paste0(Frequency, " (", Percentage, "%)"))  # Generate labels
  
  return(summary_df)
}

# Plotting function: Horizontal bar plot with gradient colors and labels
plot_summary_horizontal <- function(summary_df, plot_title) {
  ggplot(summary_df, aes(x = Frequency, y = reorder(Attribute, Frequency), fill = Frequency)) +
    geom_bar(stat = "identity", color = "black", width = 0.8) +  # Bars with black borders
    geom_text(aes(label = Label), 
              hjust = -0.1, color = "black", size = 5, fontface = "bold") +  # Label with counts and percentages
    scale_fill_gradient(low = "#D6EAF8", high = "#1B4F72") +  # Gradient fill
    theme_minimal(base_size = 18) +  # Larger font for better readability
    labs(title = plot_title, x = "Frequency", y = NULL) +  # Add clean title and x-axis label
    theme(
      legend.position = "none",  # Remove legend
      plot.title = element_text(size = 20, face = "bold", hjust = 0.5),  # Centered title
      axis.title.x = element_text(size = 16, face = "bold"),  # X-axis label size
      axis.text.x = element_text(size = 14, color = "black", face = "bold"),  # X-axis ticks
      axis.text.y = element_text(size = 14, color = "black", face = "bold"),  # Y-axis ticks
      panel.grid.major = element_line(color = "lightgray"),  # Subtle grid lines
      panel.grid.minor = element_blank()  # No minor grid lines
    ) +
    xlim(0, max(summary_df$Frequency) * 1.2)  # Padding for labels outside bars
}

# Number of participants in the dataset
total_participants <- nrow(soapbox_data)  # Replace 'soapbox_data' with your dataset

# List of attribute columns for scientist characteristics
attribute_columns <- c("Scientist_Age_MiddleAged", "Scientist_Age_Older", 
                       "Scientist_Ethnicity_White", "Scientist_Ethnicity_Black", 
                       "Scientist_Ethnicity_Asian", "Scientist_Gender_Male", 
                       "Scientist_Gender_Female", "Scientist_Dressing_Casual", 
                       "Scientist_Dressing_Formal", "Scientist_Expression_Positive", 
                       "Scientist_Expression_Neutral")

# Step 1: Prepare the summary data frame
summary_df <- prepare_summary_df(soapbox_data, attribute_columns, total_participants)

# Step 2: Print the summary table
print(summary_df)

# Step 3: Find the attribute chosen the least
least_chosen_attribute <- summary_df %>% 
  filter(Frequency == min(Frequency)) %>%
  select(Attribute)
cat("The attribute chosen the least is:", least_chosen_attribute$Attribute, "\n")

# Step 4: Save the plot with improved format
png("Scientist_Attributes_Distribution_Horizontal.png", width = 1400, height = 1000, res = 150)
plot_summary_horizontal(summary_df, "Distribution of Scientist Attributes")
dev.off()

```

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(tibble)

# Data preparation function: Summarizes attribute frequencies and calculates percentages
prepare_summary_df <- function(data, attribute_columns, total_participants) {
  # Calculate total frequency of each attribute
  summary_stats <- colSums(data[, attribute_columns], na.rm = TRUE)
  
  # Convert summary stats to a clean data frame
  summary_df <- as.data.frame(summary_stats) %>%
    rownames_to_column(var = "Attribute") %>%
    rename(Frequency = summary_stats) %>%
    mutate(Percentage = round((Frequency / total_participants) * 100, 1),
           Label = paste0(Frequency, " (", Percentage, "%)"))  # Generate labels
  
  return(summary_df)
}

# Plotting function: Horizontal bar plot with gradient colors and labels
plot_summary_horizontal <- function(summary_df, plot_title) {
  ggplot(summary_df, aes(x = Frequency, y = reorder(Attribute, Frequency), fill = Frequency)) +
    geom_bar(stat = "identity", color = "black", width = 0.8) +  # Bars with black borders
    geom_text(aes(label = Label), 
              hjust = -0.1, color = "black", size = 5, fontface = "bold") +  # Label with counts and percentages
    scale_fill_gradient(low = "#D6EAF8", high = "#1B4F72") +  # Gradient fill
    theme_minimal(base_size = 18) +  # Larger font for better readability
    labs(title = plot_title, x = "Frequency", y = NULL) +  # Add clean title and x-axis label
    theme(
      legend.position = "none",  # Remove legend
      plot.title = element_text(size = 20, face = "bold", hjust = 0.5),  # Centered title
      axis.title.x = element_text(size = 16, face = "bold"),  # X-axis label size
      axis.text.x = element_text(size = 14, color = "black", face = "bold"),  # X-axis ticks
      axis.text.y = element_text(size = 14, color = "black", face = "bold"),  # Y-axis ticks
      panel.grid.major = element_line(color = "lightgray"),  # Subtle grid lines
      panel.grid.minor = element_blank()  # No minor grid lines
    ) +
    xlim(0, max(summary_df$Frequency) * 1.2)  # Padding for labels outside bars
}

# Number of participants in the dataset
total_participants <- nrow(soapbox_data)  # Replace 'soapbox_data' with your dataset

# List of attribute columns for scientist characteristics
attribute_columns <- c("Scientist_Age_MiddleAged", "Scientist_Age_Older", 
                       "Scientist_Ethnicity_White", "Scientist_Ethnicity_Black", 
                       "Scientist_Ethnicity_Asian", "Scientist_Gender_Male", 
                       "Scientist_Gender_Female", "Scientist_Dressing_Casual", 
                       "Scientist_Dressing_Formal", "Scientist_Expression_Positive", 
                       "Scientist_Expression_Neutral")

# Step 1: Prepare the summary data frame
summary_df <- prepare_summary_df(soapbox_data, attribute_columns, total_participants)

# Step 2: Print the summary table
print(summary_df)

# Step 3: Find the attribute chosen the least
least_chosen_attribute <- summary_df %>% 
  filter(Frequency == min(Frequency)) %>%
  select(Attribute)
cat("The attribute chosen the least is:", least_chosen_attribute$Attribute, "\n")

# Step 4: Save the plot with improved format
png("Scientist_Attributes_Distribution_Horizontal.png", width = 1400, height = 1000, res = 150)
plot_summary_horizontal(summary_df, "Distribution of Scientist Attributes")
dev.off()

```


```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(tibble)

# Data preparation function: Summarizes attribute frequencies and calculates percentages
prepare_summary_df <- function(data, attribute_columns, total_participants) {
  # Calculate total frequency of each attribute
  summary_stats <- colSums(data[, attribute_columns], na.rm = TRUE)
  
  # Convert summary stats to a clean data frame
  summary_df <- as.data.frame(summary_stats) %>%
    rownames_to_column(var = "Attribute") %>%
    rename(Frequency = summary_stats) %>%
    mutate(Percentage = round((Frequency / total_participants) * 100, 1),
           Label = paste0(Frequency, " (", Percentage, "%)"))  # Generate labels
  
  return(summary_df)
}

# Plotting function: Horizontal bar plot with dynamic label adjustments
plot_summary_horizontal <- function(summary_df, plot_title) {
  ggplot(summary_df, aes(x = Frequency, y = reorder(Attribute, Frequency), fill = Frequency)) +
    geom_bar(stat = "identity", color = "black", width = 0.8) +  # Bars with black borders
    
    # Dynamic label position and color for narrow bars
    geom_text(aes(label = Label, 
                  hjust = ifelse(Frequency / max(Frequency) > 0.3, 1.1, -0.1)),  # Inside for wide, outside for narrow
              color = ifelse(summary_df$Frequency / max(summary_df$Frequency) > 0.3, "white", "black"), 
              size = 5, fontface = "bold") +
    
    scale_fill_gradient(low = "#D6EAF8", high = "#1B4F72") +  # Gradient fill
    theme_minimal(base_size = 18) +  # Larger font for better readability
    labs(title = plot_title, x = "Frequency", y = NULL) +  # Add clean title and x-axis label
    theme(
      legend.position = "none",  # Remove legend
      plot.title = element_text(size = 20, face = "bold", hjust = 0.5),  # Centered title
      axis.title.x = element_text(size = 16, face = "bold"),  # X-axis label size
      axis.text.x = element_text(size = 14, color = "black", face = "bold"),  # X-axis ticks
      axis.text.y = element_text(size = 14, color = "black", face = "bold"),  # Y-axis ticks
      panel.grid.major = element_line(color = "lightgray"),  # Subtle grid lines
      panel.grid.minor = element_blank()  # Remove minor grid lines
    ) +
    xlim(0, max(summary_df$Frequency) * 1.3)  # Extra padding for outside labels
}

# Number of participants in the dataset
total_participants <- nrow(soapbox_data)  # Replace 'soapbox_data' with your dataset

# List of attribute columns for scientist characteristics
attribute_columns <- c("Scientist_Age_MiddleAged", "Scientist_Age_Older", 
                       "Scientist_Ethnicity_White", "Scientist_Ethnicity_Black", 
                       "Scientist_Ethnicity_Asian", "Scientist_Gender_Male", 
                       "Scientist_Gender_Female", "Scientist_Dressing_Casual", 
                       "Scientist_Dressing_Formal", "Scientist_Expression_Positive", 
                       "Scientist_Expression_Neutral")

# Step 1: Prepare the summary data frame
summary_df <- prepare_summary_df(soapbox_data, attribute_columns, total_participants)

# Step 2: Print the summary table
print(summary_df)

# Step 3: Find the attribute chosen the least
least_chosen_attribute <- summary_df %>% 
  filter(Frequency == min(Frequency)) %>%
  select(Attribute)
cat("The attribute chosen the least is:", least_chosen_attribute$Attribute, "\n")

# Step 4: Save the plot with improved format
png("Scientist_Attributes_Distribution_Horizontal.png", width = 1400, height = 1000, res = 150)
plot_summary_horizontal(summary_df, "Distribution of Scientist Attributes")
dev.off()

```

```{r,eval=FALSE}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(tibble)

# Set up the data preparation
prepare_summary_df <- function(data, attribute_columns, total_participants) {
  # Calculate frequency of each attribute
  summary_stats <- colSums(data[, attribute_columns])
  
  # Convert summary stats to a data frame
  summary_df <- as.data.frame(summary_stats) %>%
    rownames_to_column(var = "Attribute") %>%
    rename(Frequency = summary_stats) %>%
    mutate(Percentage = round((Frequency / total_participants) * 100, 1))  # Calculate percentage for each attribute
  
  return(summary_df)
}

# Plotting function for the summary statistics in horizontal format with gradient colors
plot_summary_horizontal <- function(summary_df) {
  ggplot(summary_df, aes(x = Frequency, y = reorder(Attribute, Frequency), fill = Frequency)) +
    geom_bar(stat = "identity", color = "black") +  # Bars with a black border
    geom_text(aes(label = paste(Frequency, " (", Percentage, "%)", sep = "")), 
              hjust = -0.1, color = "black", size = 5, fontface = "bold") +  # Adjust text size and position
    scale_fill_gradient(low = "lightblue", high = "darkblue") +  # Gradient from light blue to dark blue
    theme_minimal(base_size = 16, base_family = "Helvetica") +  # Larger font size for readability
    labs(title = NULL, x = "Frequency", y = NULL) +  # No title, labels for clarity
    theme(
      legend.position = "none",  # Remove the legend
      axis.text.x = element_text(size = 14, color = "black", face = "bold"),  # Adjust x-axis text
      axis.text.y = element_text(size = 14, color = "black", face = "bold"),  # Adjust y-axis text
      plot.background = element_rect(fill = "white", color = NA),  # White background
      panel.background = element_rect(fill = "white", color = NA),  # White panel
      panel.grid.major = element_line(color = "gray"),  # Gray grid lines
      panel.grid.minor = element_blank()  # No minor grid lines
    ) +
    xlim(0, max(summary_df$Frequency) * 1.1)  # Adjust x-axis to fit text inside the plot
}


# Replace this with the actual number of participants
total_participants <- nrow(soapbox_data)

# List of attribute columns for scientist characteristics
attribute_columns <- c("Scientist_Age_MiddleAged", "Scientist_Age_Older", 
                       "Scientist_Ethnicity_White", "Scientist_Ethnicity_Black", 
                       "Scientist_Ethnicity_Asian", "Scientist_Gender_Male", 
                       "Scientist_Gender_Female", "Scientist_Dressing_Casual", 
                       "Scientist_Dressing_Formal", "Scientist_Expression_Positive", 
                       "Scientist_Expression_Neutral")

# Prepare the summary data frame
summary_df <- prepare_summary_df(soapbox_data, attribute_columns, total_participants)

# Display the table
print(summary_df)

# Find the attribute chosen the least
least_chosen_attribute <- summary_df %>% 
  filter(Frequency == min(Frequency)) %>%
  select(Attribute)
cat("The attribute chosen the least is:", least_chosen_attribute$Attribute, "\n")

# Save the plot as a PNG file with horizontal format and gradient colors
png("Scientist_Attributes_Distribution_Horizontal.png", width = 1400, height = 1000, res = 150)  # Adjust dimensions
plot_summary_horizontal(summary_df)  # Generate the plot
dev.off()  # Close the device

```





```{r}

soapbox_data_categorised <- soapbox_data
# Define the categorical conversion for each relevant column
convert_to_numeric <- function(column, levels) {
  factor_column <- factor(column, levels = levels, labels = 1:length(levels))
  as.numeric(factor_column)
}

# List of levels in order for each column (example)
levels_list <- list(
  Improves_Lives = c("Completely disagree",  "Completely agree"),
  Good_With_Scientific_Method = c("Completely disagree", "Somewhat disagree",  "Somewhat agree", "Completely agree"),
  Consider_Career_In_STEM = c("Completely disagree", "Somewhat disagree",  "Somewhat agree", "Completely agree"),
  Inspired_By_Women_NonBinary_Scientists = c( "Somewhat agree", "Completely agree"),
  Interest_Increased_By_Event = c("Somewhat agree", "Completely agree"),
  Interest_In_Science_Research = c( "Somewhat interested", "Interested", "Very interested")
)

# Apply the conversion function
soapbox_data_categorised$Improves_Lives <- convert_to_numeric(soapbox_data_categorised$Improves_Lives, levels_list$Improves_Lives)
soapbox_data_categorised$Good_With_Scientific_Method <- convert_to_numeric(soapbox_data_categorised$Good_With_Scientific_Method, levels_list$Good_With_Scientific_Method)
soapbox_data_categorised$Consider_Career_In_STEM <- convert_to_numeric(soapbox_data_categorised$Consider_Career_In_STEM, levels_list$Consider_Career_In_STEM)
soapbox_data_categorised$Inspired_By_Women_NonBinary_Scientists <- convert_to_numeric(soapbox_data_categorised$Inspired_By_Women_NonBinary_Scientists, levels_list$Inspired_By_Women_NonBinary_Scientists)
soapbox_data_categorised$Interest_Increased_By_Event <- convert_to_numeric(soapbox_data_categorised$Interest_Increased_By_Event, levels_list$Interest_Increased_By_Event)
soapbox_data_categorised$Interest_In_Science_Research <- convert_to_numeric(soapbox_data_categorised$Interest_In_Science_Research, levels_list$Interest_In_Science_Research)

# Check the structure of the data to verify the categorical variables
str(soapbox_data_categorised)
View(soapbox_data_categorised)


```



```{r}


# Load necessary libraries
library(dplyr)

# Convert 'Age' column to numeric categories
soapbox_data_categorised <- soapbox_data_categorised %>%
  mutate(
    Age = case_when(
      Age == "12-15" ~ 1,
      Age == "16-20" ~ 2,
      Age == "21-30" ~ 3,
      Age == "31-40" ~ 4,
      Age == "41-50" ~ 5,
      Age == "51-60" ~ 6,
      Age =="Older than 61" ~ 7,
      Age == "Prefer not to answer" ~ 8,
      TRUE ~ NA_real_  # Handle missing values or other cases
    )
  )

# Convert 'Gender' column to numeric categories
soapbox_data_categorised <- soapbox_data_categorised %>%
  mutate(
    Gender = case_when(
      Gender == "Male" ~ 1,
      Gender == "Female" ~ 2,
      Gender == "Non-binary" ~ 3,
      Gender == "Prefer not to answer" ~ 4,
      TRUE ~ NA_real_  # Handle missing values or other cases
    )
  )

# Convert 'Educational_Background' column to numeric categories
soapbox_data_categorised <- soapbox_data_categorised %>%
  mutate(
    Educational_Background = case_when(
      Educational_Background == "Elementary School" ~ 1,
      Educational_Background == "High school" ~ 2,
      Educational_Background == "College Diploma" ~ 3,
      Educational_Background == "Undergraduate degree" ~ 4,
      Educational_Background == "Graduate degree" ~ 5,
      Educational_Background == "Red Seal" ~ 6,
      TRUE ~ NA_real_  # Handle missing values or other cases
    )
  )

# View the updated data
str(soapbox_data_categorised)  # Check the structure of the data to verify the changes
head(soapbox_data_categorised)  # Display the first few rows to verify the changes
View(soapbox_data_categorised)  # View the data in RStudio viewer
# Save the data frame as an Excel file
write.xlsx(soapbox_data_categorised, "soapbox_data_categorised.xlsx", rowNames = FALSE)


```


```{r}
# Read the saved Excel file back into R
data <- read.xlsx("soapbox_data_categorised.xlsx")

# View the data to verify
head(data)
str(data)
View(data)


```






```{r}
# Load necessary libraries
library(dplyr)

# Check for columns that have all zero entries
zero_columns <- sapply(data, function(col) all(col == 0))

# Display the columns that have all-zero entries (if any)
if (any(zero_columns)) {
  print(names(data)[zero_columns])
  # Remove columns that have all zero entries
  data <- data[, !zero_columns]
  cat("Columns with all-zero entries have been removed.\n")
} else {
  cat("No columns with all-zero entries found.\n")
}

# Display the structure of the cleaned data
str(data)

```





```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(tibble)

# Calculate the frequency of each ethnicity response
ethnicity_sums <- colSums(data[, grepl("Response_Ethnicity", names(data))])

# Assuming total number of participants
total_participants <- nrow(data)

# Convert to data frame for plotting
ethnicity_df <- as.data.frame(ethnicity_sums) %>%
  rownames_to_column(var = "Ethnicity") %>%
  rename(Frequency = ethnicity_sums) %>%
  mutate(Percentage = round((Frequency / total_participants) * 100, 1),
         Label = paste0(Frequency, " (", Percentage, "%)"))  # Add labels with counts and percentages

# Dynamic bar label adjustments (position & color based on bar width)
plot <- ggplot(ethnicity_df, aes(x = Frequency, y = reorder(Ethnicity, Frequency), fill = Frequency)) +
  geom_bar(stat = "identity", color = "black", width = 0.8) +  # Bars with black borders
  
  # Dynamic labels: position and color adjustments
  geom_text(aes(label = Label, 
                hjust = ifelse(Frequency / max(Frequency) > 0.3, 1.1, -0.2),  # Inside for wide, outside for narrow
                color = ifelse(Frequency / max(Frequency) > 0.3, "white", "black")),  # Text color logic
            size = 5, fontface = "bold") +
  
  scale_fill_gradient(low = "lightblue", high = "darkblue") +  # Gradient from light to dark blue
  scale_color_identity() +  # Allows manual control of text color
  
  # Improve aesthetics and readability
  theme_minimal(base_size = 14, base_family = "Helvetica") +
  labs(title = "Ethnicity Distribution", x = "Frequency", y = NULL) +
  theme(
    legend.position = "none",  # Remove legend
    axis.text.x = element_text(size = 12, color = "black", face = "bold"),  # X-axis text size
    axis.text.y = element_text(size = 12, color = "black", face = "bold"),  # Y-axis text size
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),  # Title style
    panel.grid.major = element_line(color = "gray"),  # Subtle grid lines
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
  xlim(0, max(ethnicity_df$Frequency) * 1.3)  # Extra padding for outside labels

# Save the horizontal plot as a JPG file
jpeg("Ethnicity_Distribution_Horizontal.jpg", width = 1200, height = 900, res = 150)
print(plot)  # Generate and display the updated plot
dev.off()  # Close the device

# Display the table of counts and percentages
print(ethnicity_df)

```

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(tibble)

# Calculate the frequency of each ethnicity response
ethnicity_sums <- colSums(data[, grepl("Response_Ethnicity", names(data))])

# Assuming total number of participants is 30
total_participants <- nrow(data)

# Convert to data frame for plotting
ethnicity_df <- as.data.frame(ethnicity_sums)
ethnicity_df <- ethnicity_df %>%
  rownames_to_column(var = "Ethnicity") %>%
  rename(Frequency = ethnicity_sums) %>%
  mutate(Percentage = round((Frequency / total_participants) * 100, 1))  # Calculate percentage based on participants

# Create a horizontal bar plot with gradient colors and improved readability
plot <- ggplot(ethnicity_df, aes(x = Frequency, y = reorder(Ethnicity, Frequency), fill = Frequency)) +
  geom_bar(stat = "identity", color = "black") +  # Gradient fill with black border
  geom_text(aes(label = paste0(Frequency, " (", Percentage, "%)")), 
            hjust = -0.2, color = "black", size = 3, fontface = "bold") +  # Position labels outside bars
  scale_fill_gradient(low = "lightblue", high = "darkblue") +  # Gradient color from light to dark blue
  theme_minimal(base_size = 14, base_family = "Helvetica") +  # White background with readable font
  labs(title = NULL, x = "Frequency", y = NULL) +  # No title, axis labels for clarity
  theme(
    legend.position = "none",  # Remove the legend
    axis.text.x = element_text(size = 12, color = "black", face = "bold"),  # Adjust x-axis text size and style
    axis.text.y = element_text(size = 12, color = "black", face = "bold"),  # Adjust y-axis text size and style
    plot.background = element_rect(fill = "white", color = NA),  # White background
    panel.background = element_rect(fill = "white", color = NA),  # White panel background
    panel.grid.major = element_line(color = "gray"),  # Gray grid lines
    panel.grid.minor = element_blank()  # No minor grid lines
  ) +
  xlim(0, max(ethnicity_df$Frequency) * 1.1)  # Adjust x-axis limits to fit labels

# Save the horizontal plot as a JPG file
jpeg("Ethnicity_Distribution_Horizontal.jpg", width = 1200, height = 900, res = 150)
print(plot)  # Generate and display the updated plot
dev.off()  # Close the device

# Display the table of counts and percentages
print(ethnicity_df)

```




```{r}
# Load required libraries
library(dplyr)
library(ggplot2)
library(tidyr)

# 1. Analyze ethnicity selection by White respondents
ethnicity_by_white <- data %>%
  filter(Response_Ethnicity_White == 1) %>%  # Only White respondents
  summarise(
    White = sum(Scientist_Ethnicity_White, na.rm = TRUE),
    Black = sum(Scientist_Ethnicity_Black, na.rm = TRUE),
    Asian = sum(Scientist_Ethnicity_Asian, na.rm = TRUE)
  ) %>%
  gather(key = "Ethnicity", value = "Count") %>%
  mutate(
    Total_Respondents = nrow(data %>% filter(Response_Ethnicity_White == 1)),
    Percentage = round((Count / Total_Respondents) * 100, 1),
    Label = paste0(Count, " (", Percentage, "%)")
  ) %>%
  arrange(desc(Percentage))

# 2. Analyze gender selection by female respondents
gender_by_female <- data %>%
  filter(Gender == 2) %>%  # Only Female respondents
  summarise(
    Male = sum(Scientist_Gender_Male, na.rm = TRUE),
    Female = sum(Scientist_Gender_Female, na.rm = TRUE)
  ) %>%
  gather(key = "Gender", value = "Count") %>%
  mutate(
    Total_Respondents = nrow(data %>% filter(Gender == 2)),
    Percentage = round((Count / Total_Respondents) * 100, 1),
    Label = paste0(Count, " (", Percentage, "%)")
  ) %>%
  arrange(desc(Percentage))

# Function for dynamic labels and readability
plot_dynamic_labels <- function(data, x_label, y_label, title) {
  ggplot(data, aes(x = Count, y = reorder(!!sym(x_label), Count), fill = Percentage)) +
    geom_bar(stat = "identity", color = "black", width = 0.8) +
    geom_text(aes(label = Label, 
                  hjust = ifelse(Count / max(Count) > 0.3, 1.1, -0.2), 
                  color = ifelse(Count / max(Count) > 0.3, "white", "black")),
              size = 5, fontface = "bold") +
    scale_fill_gradient(low = "lightblue", high = "darkblue") +
    scale_color_identity() +
    labs(title = title, x = "Count", y = NULL) +
    theme_minimal(base_size = 16) +
    theme(
      plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
      axis.text.x = element_text(size = 14, face = "bold", color = "black"),
      axis.text.y = element_text(size = 14, face = "bold", color = "black"),
      legend.position = "none",
      panel.grid.major = element_line(color = "gray"),
      panel.grid.minor = element_blank()
    ) +
    xlim(0, max(data$Count) * 1.2)  # Extra space for outside labels
}

# 3. Plot ethnicity selection by White respondents
plot_ethnicity <- plot_dynamic_labels(ethnicity_by_white, "Ethnicity", "Count", 
                                      "Ethnicity Selection by White Respondents")
ggsave("ethnicity_selection_by_white_with_labels.png", plot_ethnicity, width = 8, height = 6, dpi = 150)

# 4. Plot gender selection by female respondents
plot_gender <- plot_dynamic_labels(gender_by_female, "Gender", "Count", 
                                   "Gender Selection by Female Respondents")
ggsave("gender_selection_by_female_with_labels.png", plot_gender, width = 8, height = 6, dpi = 150)

# Display results
print("Ethnicity selection by White respondents:")
print(ethnicity_by_white)

print("Gender selection by Female respondents:")
print(gender_by_female)

```


```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(tidyr)   # For pivot_longer
library(ggalluvial)  # For alluvial plots

# ------------------ Proportional Bar Plot Preparation ------------------
# Data Preparation: Proportional Distribution
proportional_data <- data %>%
  pivot_longer(cols = starts_with("Scientist_"), names_to = "Attribute", values_to = "Selected") %>%
  filter(Selected == 1) %>%
  pivot_longer(cols = starts_with("Response_Ethnicity"), names_to = "Ethnicity", values_to = "Ethnicity_Selected") %>%
  filter(Ethnicity_Selected == 1) %>%
  group_by(Ethnicity, Attribute) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(
    Total = sum(Count),
    Proportion = round(Count / Total, 3),
    Label = paste0(round(Proportion * 100, 1), "%")  # Add percentage labels
  )

# Plot: Proportional Bar Plot with Dynamic Labels
proportional_plot <- ggplot(proportional_data, aes(x = Ethnicity, y = Proportion, fill = Attribute)) +
  geom_bar(stat = "identity", position = "stack", color = "black") +  # Black border for clarity
  geom_text(aes(label = Label),
            position = position_stack(vjust = 0.5),  # Position labels dynamically
            size = 4, fontface = "bold", color = "black") +
  scale_fill_brewer(palette = "Set3") +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  ) +
  labs(
    title = "Proportional Distribution of Scientist Attributes by Ethnicity",
    x = "Ethnicity",
    y = "Proportion",
    fill = "Scientist Attributes"
  )

# Save the Proportional Plot
ggsave("proportional_distribution_by_ethnicity.png", proportional_plot, width = 10, height = 7, dpi = 300)

# ------------------ Alluvial Plot Preparation ------------------
# Data Preparation for Alluvial Plot
alluvial_data <- proportional_data %>%
  select(Ethnicity, Attribute, Count)

# Plot: Alluvial Diagram with Improved Label Positioning
alluvial_plot <- ggplot(alluvial_data,
       aes(axis1 = Ethnicity, axis2 = Attribute, y = Count)) +
  geom_alluvium(aes(fill = Attribute), color = "darkgray", alpha = 0.8) +
  geom_stratum(color = "black", size = 0.5) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
            size = 3.5, fontface = "bold") +
  scale_x_discrete(limits = c("Ethnicity", "Attribute"), expand = c(0.15, 0.05)) +
  scale_fill_brewer(palette = "Set3") +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  ) +
  labs(
    title = "Alluvial Plot of Ethnicity and Scientist Attribute Relationships",
    x = "", y = "Count",
    fill = "Scientist Attributes"
  )

# Save the Alluvial Plot
ggsave("enhanced_alluvial_plot.png", alluvial_plot, width = 12, height = 8, dpi = 300)

# ------------------ Output Results ------------------
print("Proportional Data Table:")
print(proportional_data)


```




